Performance comparison of Dask and Apache Spark on HPC systems

### 2022-02-19 latex/paper.tex
[ADD] Highligh modified parts.

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~\cite{OpenSFS-lustre}, offering parallel file accesses.

@@

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences
and similarities, for HPC, between Dask and Spark for data-intensive applications.

@@

\section{Background}
\subsection{Lustre}
The Lustre file system is a parallel file system of reference in HPC environments.
Lustre offers a POSIX-compliant interface with the file system.
Installed in some of the most powerful supercomputers, the design of Lustre allows
scaling to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.

@@

While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
HDFS makes use of permanent local storage on compute node, which is usually not feasable in HPC.
Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.

@@

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
All of the compute and storage nodes are inter-connected by a \SI{10}{\giga\bit\per\second} Ethernet link through a switch.
The login node, hosting an NFS, is also connected by a \SI{10}{\giga\bit\per\second} Ethernet link to the compute nodes.

@@

Our system uses similar hardware as Compute Canada Beluga, and replicates as
close as possible it's configuration.
This let us make a fair comparison of Spark and Dask in the condition of a realistic
HPC system.

@@

\subsection{Applications}
In this section, we describe the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}\cite{gorgolewski2017bids}.

@@

The scripts used for our experiements and the results generated are available on
GitHub at: \href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}.

@@

\section{Discussion}
\subsection{I/O bottleneck}
We used Lustre as shared file system as it is a defacto solution in HPC systems.
The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
This suggests that increasing the parallelism level of the file system, including through potential configuration optimizations, might help alleviate the I/O bottleneck observed.

@@

These results suggest that increasing the parallelism of the file system is critical when scaling the compute resources to process a data-intensive application.
Otherwise, the compute resource might be underutilized.

@@

\subsection{Memory management}
Our Kmeans experiment shows a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
That is, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
For our experiments, we allocated all available memory on the compute nodes.
Therefore, it was not possible to allocate more resources to test this hypothesis.

>>>>

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\HL}[1]{\hl{#1}}

@@

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
\HL{Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.}

@@

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences
and similarities\HL{, for HPC,} between Dask and Spark for data-intensive applications.

@@

\section{Background}
\subsection{Lustre}
The Lustre file system is a parallel file system \HL{of reference in} HPC environments.
Lustre offers a POSIX-compliant interface with the file system.
Installed in some of the most powerful supercomputers, the design of Lustre allows
scaling to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.

@@

\HL{While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
	HDFS makes use of permanent local storage on compute node, which is usually not feasable in HPC.
Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.}

@@

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a switch.
	The login node, hosting an NFS, is also connected by a {\SI{10}{\giga\bit\per\second}} Ethernet link to the compute nodes.}

@@

\HL{Our system uses similar hardware as Compute Canada Beluga, and replicates as
	close as possible it's configuration.
	This let us make a fair comparison of Spark and Dask in the condition of a realistic
HPC system.}

@@

\subsection{Applications}
In this section, we describe the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}\HL{{\cite{gorgolewski2017bids}}}.

@@

\HL{The scripts used for our experiements and the results generated are available on
	GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.}

@@

\section{Discussion}
\subsection{I/O bottleneck}
We used Lustre as shared file system as it is a \HL{de facto} solution in HPC systems.
The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
This suggests that increasing the parallelism level of the file system, including through potential configuration optimizations, might help alleviate the I/O bottleneck observed.

@@

These results suggest that increasing the parallelism of the file system is critical when scaling the compute resources to process a data-intensive application.
Otherwise, the compute resource might be underutilized.
% TODO talk about the limitation of our current setup.

@@

\subsection{Memory management}
Our Kmeans experiment shows a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and \HL{are allocated} the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
\HL{That is,} with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
\HL{For our experiments, we allocated all available memory on the compute nodes.
Therefore, it was not possible to allocate more resources to test this hypothesis.}
